%\documentclass[manuscript]{biometrika}
\documentclass[lineno]{biometrika}

\usepackage{amsmath}
%\usepackage{varwidth}

%% Please use the following statements for
%% managing the text and math fonts for your papers:
\usepackage{times}
\usepackage{bm}
\usepackage{natbib}

\makeatletter

\usepackage{verbatim}
\usepackage{graphicx,epsfig}
\usepackage{amsmath,amssymb,latexsym, amsfonts, amscd}

\usepackage[plain,noend]{algorithm2e}

\makeatletter
\renewcommand{\algocf@captiontext}[2]{#1\algocf@typo. \AlCapFnt{}#2} % text of caption
\renewcommand{\AlTitleFnt}[1]{#1\unskip}% default definition
\def\@algocf@capt@plain{top}
\renewcommand{\algocf@makecaption}[2]{%
  \addtolength{\hsize}{\algomargin}%
  \sbox\@tempboxa{\algocf@captiontext{#1}{#2}}%
  \ifdim\wd\@tempboxa >\hsize%     % if caption is longer than a line
    \hskip .5\algomargin%
    \parbox[t]{\hsize}{\algocf@captiontext{#1}{#2}}% then caption is not centered
  \else%
    \global\@minipagefalse%
    \hbox to\hsize{\box\@tempboxa}% else caption is centered
  \fi%
  \addtolength{\hsize}{-\algomargin}%
}
\makeatother
\SetAlCapSty{}
\def\Bka{{\it Biometrika}}
\def\AIC{\textsc{aic}}
\def\T{{ \mathrm{\scriptscriptstyle T} }}
\def\v{{\varepsilon}}


\newcommand{\figref}[1]{\figurename~\ref{#1}}
\usepackage[usenames]{color}

\def\cM{\mathcal{M}}
\def\cL{\mathcal{L}}
\def\cH{\mathcal{H}}

\def\cO{\mathcal{O}}
\def\cT{\mathcal{T}}
\def\cU{\mathcal{U}}
\def\cA{\mathcal{A}}
\def\cV{\mathcal{V}}
\def\cD{\mathcal{D}}
\def\cG{\mathcal{G}}
\def\cW{\mathcal{W}}
\def\etr{\text{etr}}
\def\hZ{\hat{Z}}
\def\hX{\hat{X}}
\DeclareMathOperator{\Tr}{Trace}
\DeclareMathOperator{\diag}{Diag}
\def\bkappa{\boldsymbol\kappa}
\def\bkappa{\kappa}

\def\bY{\mathbf{Y}}
\def\cY{\mathcal{Y}}
\def\cX{\mathcal{X}}
\def\tD{\tilde{D}}
\def\bX{\mathbb{X}}
\def\bU{\mathbb{U}}
\def\htheta{\hat{\theta}}
\newcommand{\dif}{\mathrm{d}}

\def\ml{{p}_{\text{ML}}}
\def\seq{{p}_{\text{seq}}}
%\allowdisplaybreaks

\begin{document}

\jname{Biometrika}
%% The year, volume, and number are determined on publication
\jyear{2014}
\jvol{99}
\jnum{1}
%% The \doi{...} and \accessdate commands are used by the production team
%\doi{10.1093/biomet/asm023}
\accessdate{Advance Access publication on 31 July 2014}
\copyrightinfo{\Copyright\ 2014 Biometrika Trust\goodbreak {\em Printed in Great Britain}}

%% These dates are usually set by the production team
\received{June 2014}
\revised{June 2014}

%% The left and right page headers are defined here:
%\markboth{A. C. Davison, R. Gessner \and D. M. Titterington}{Biometrika style}

%% Here are the title, author names and addresses
\markboth{Vinayak Rao, Lizhen Lin, \and David Dunson}{Data augmentation for models based on rejection sampling}
\title{Data augmentation for models based on rejection sampling}

\author{Vinayak Rao}
\affil{Department of Statistics, Purdue University, West Lafayette, Indiana 47907, U.S.A. \email{varao@purdue.edu} }
\author{Lizhen Lin}
\affil{Department of Statistics and Data Science, the University of Texas, Austin, Texas 78712, U.S.A.  \email{lizhen.lin@austin.utexas.edu}}
\author{David Dunson}
\affil{Department of Statistical Science, Duke University, Durham, North Carolina 27708, U.S.A. \email{dunson@duke.edu}}


\maketitle

\begin{abstract}
We present a data augmentation scheme to perform Markov chain Monte Carlo inference for models where data generation involves a
rejection sampling algorithm. 
Our idea %, which seems to be missing in the literature, is %to instantiate the rejected proposals preceding each data point, and we show that this can be done 
is a simple scheme to instantiate the rejected proposals preceding each data point. %easily and efficiently.
The resulting joint probability over observed and rejected variables can be much simpler than the marginal distribution over the
observed variables, which often involves intractable integrals. 
%Our algorithm is an instance of a growing body of work on exact Markov chain Monte Carlo inference for 
%doubly-intractable distributions and 
We consider three problems: modeling flow-cytometry measurements subject to truncation; 
the Bayesian analysis of the matrix Langevin distribution on the Stiefel manifold; and
Bayesian inference for a nonparametric Gaussian process density model.
The latter two are instances of doubly-intractable Markov chain Monte Carlo problems, where evaluating the likelihood 
is intractable. % distributions and we consider two such problems. 
Our experiments demonstrate superior performance over state-of-the-art sampling algorithms for such problems.

~\\
\noindent \emph{Some key words}: Bayesian inference; Density estimation;  Intractable likelihood; Gaussian process; Matrix Langevin distribution; Markov Chain Monte Carlo; Rejection sampling; Truncation.

\end{abstract}



\section{Introduction}


\input{intro.tex}


\section{Rejection sampling}
\input{rejection.tex}

\section{Flow cytometry data} \label{sec:flow}
\input{flowcyto.tex}

\section{Bayesian inference for the matrix Langevin distribution} \label{sec:Bays_inf}
\input{bayes_model.tex}


%\section{Experiments}  \label{sec:Bayes_expt}
\input{bayes_expts.tex}

\section{The Gaussian process density sampler}\label{sec:gpds}
\input{gpds.tex}


\section{Future work}\label{sec:conc}
\input{conclusion.tex}


\section*{Acknowledgement}
This work was supported by the National Institute of Environmental Health Sciences of the National Institute of Health.
We are grateful to the editor and reviewers for valuable comments.


\appendix
\section{Proofs}
\input{proofs.tex} \label{sec:proofs}
%\input{proofs_long.tex} 

\section{Gradient information}
\input{gradient.tex} \label{sec:gradient}


\bibliographystyle{biometrika}
%\bibliographystyle{plain}
\bibliography{refs}

\end{document}
